
---
title: "Data cleaning for the Price prediction of apartments in Barcelona"
# subtitle: "Data cleaning for the Price prediction of apartments in Barcelona"
author: "Viktória Kónya"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=1.8cm
fontsize: 9pt
output:

  prettydoc::html_pretty:
    highlight: github
    toc: true
    theme: cayman
---

What the code does:

* Imports Barcelona Airbnb data from GitHub repo
* Sample selection
* Data cleaning and feature engineering
* Saves the workfile

```{r setup, include = F}

# Chunk setup
knitr::opts_chunk$set(warning = F, message = F)

```

```{r folder_library_setup, echo = F, include = F}

# Clear environment
rm(list=ls())

# Folder locations
main <- getwd()
data_raw <- paste0(main, '/data/raw')
data_work <- paste0(main, '/data/work')
output <- paste0(main, '/output')

# Import libraries
library(tidyverse)
library(skimr)
library(data.table)
library(modelsummary)

```


# Data import
```{r data_import}

# Import data - 15.707 obs, 74 vars
# listings <- read_csv("https://raw.githubusercontent.com/viktoriakonya/CEU-DA3/main/A2/data/raw/listings_20211207.csv") 
listings <- read_csv(paste0(data_raw,"/listings_20211207.csv")) 

# Convert to data table
df <- data.table(listings)

# Look at the dataset - 74 vars; 15.707 obs.
skim(listings)

```


# Sample selection

### 1. step: Filter for Entire home / Apartments only 
```{r}

df[, .(count = .N), by = .(room_type)]

# Entire home / Apartments only - 8.407 obs.
df <- df[room_type %in% c("Entire home/apt")]

```
### 2. step: Filter out irrelevant property types
```{r}

df[, .(count = .N), by = .(property_type)]

# Relevant property types only  - 8.341 obs.
df <- df[!(property_type %in% c("Boat", "Barn", "Floor", "Camper/RV", "", "Room in aparthotel", "Entire vacation home", "Entire villa"))]

```
**Note**: Later categories with few listings chould be grouped.

### 3. step: Filter for apartments with 2-6 accommodates
```{r}

df[, .(count = .N), by = .(accommodates)][order(accommodates)]

# Apartments with 2-6 accommodates - 7260 obs.
df <- df[accommodates >= 2 & accommodates <= 6]

```

### 4. step:  Filter for apartments where the number of beds <= 6 or NA
```{r}

df[, .(count = .N), by = .(beds)][order(beds)]

# Apartments with beds <= 6 - 7241 obs.
df <- df[beds <= 6 | is.na(beds)]

```
**Note**: NAs should be imputed later.

### 5. step:  Filter out apartments with 0 bathrooms
```{r}

df[,.(count = .N), by =.(bathrooms_text)] # 1 obs.

# Apartments with bathrooms > 0 - 7240 obs.
df <- df[bathrooms_text != "0 baths" | is.na(bathrooms_text)]

```
**Note**: NA should be imputed later.

### 6. step:  Filter out apartments with minimum nights > 90 days
```{r}

df[minimum_nights > 90,.(count = .N)] # 61 obs.

# Apartments with bathrooms > 0 - 7179 obs.
df <- df[minimum_nights <= 90]

```


# Drop useless fields and with high missing rate

```{r}

# Check variables with only missing values
missings <- sapply(df, function(x) sum(is.na(x)))
missings[missings > 0]
# calendar_updated, bathrooms have only missings

# Drop variables with only NA, host-related information and useless fields 
drops <- c("listing_url", "scrape_id", "last_scraped", "name", "description", "neighborhood_overview",
           "picture_url", "host_id", "host_url", "host_name", "host_since", "host_location", "host_about", "host_response_time",
           "host_response_rate", "host_acceptance_rate", "host_is_superhost", 
           "host_thumbnail_url","host_picture_url","host_neighbourhood", "host_listings_count",
           "host_total_listings_count", "host_verifications", "host_has_profile_pic", "host_identity_verified",
           "minimum_minimum_nights", "maximum_minimum_nights", "minimum_maximum_nights", "maximum_maximum_nights",
           "minimum_nights_avg_ntm", "maximum_nights_avg_ntm","availability_60", "availability_90", "availability_365",
           "number_of_reviews_ltm", "number_of_reviews_l30d", "first_review", "last_review",
           "review_scores_accuracy", "review_scores_cleanliness", "review_scores_checkin", "review_scores_communication",
           "review_scores_location", "review_scores_value", "license", "calculated_host_listings_count",
           "calculated_host_listings_count_private_rooms", "reviews_per_month",
           "bathrooms", "calendar_updated"
           )

# 24 vars, 7.240 obs
df <- df[, (drops) := NULL]

```




# Data cleaning

### 1. step: Price per night
```{r}

# Clean outcome var and rename
df$price <- gsub("\\,","",df$price) # delete commas
df$n_price <- as.numeric(gsub("\\$","",df$price)) # delete $ from the end

# Missings
datasummary_skim(df$n_price) # 0 missing

# Distribution
hist(df$n_price, breaks = 100) # Highly right-skewed, 16 extrem above 1000 USD

# Price distr. by property type
df[, .(count = .N, min_price = min(n_price), max_price = max(n_price)), by = .(property_type)]

# Distribution is highly skewed -> Take log of prices (no 0 values)
# df$n_ln_price <- log(df$n_price)

```
**Note**: Price per night is highly right-skewed with 40 listings above 600 USD and 16 extreme listings above 1000 USD. As these listings are irrelevalant for our analysis hence they were excluded them later from the analysis.

### 2. step: Bathrooms
```{r}

# Summary
df[, .(count = .N), by = .(bathrooms_text)][order(bathrooms_text)] # 3 not numeric, 1 NA

# Add number to the beginning
df$bathrooms_text_mod <- ifelse(df$bathrooms_text == "Shared half-bath", "0.5 shared half-bath", df$bathrooms_text)
df$bathrooms_text_mod <- ifelse(df$bathrooms_text == "Half-bath", "0.5 half-bath", df$bathrooms_text_mod)
df$bathrooms_text_mod <- ifelse(df$bathrooms_text == "Private half-bath", "0.5 private half-bath", df$bathrooms_text_mod)

# Create variable for number of bathroom
df$n_bathrooms <- as.numeric(sub( " .*$", "", df$bathrooms_text_mod )) 

# Check
df[, .(count = .N), by = .(bathrooms_text, bathrooms_text_mod, n_bathrooms)][order(bathrooms_text)] 
# 1 NA and 1 with 0
hist(df$n_bathrooms, breaks = 10)

# Create factor for bathroom with 3 levels: [0-0.5], [1-1.5], [2+]
df[, f_bathrooms:= cut(n_bathrooms, c(1,2,3,10), labels=c(1,2,3), right = F)  ]
df[, .(count = .N, min = min(n_bathrooms), max = max(n_bathrooms)), by = .(f_bathrooms)]


```
**Note**: The bathroom field of the dataset was not reliable, hence I used the text field to create this variable. A there are very few observations above 3 bathrooms we can create categories of the bathroom number later.


### 3. step: Amenities
```{r}

# Clean list
df$amenities <- gsub("\\{","",df$amenities)
df$amenities <- gsub("\\}","",df$amenities)
df$amenities <- gsub('\\"',"",df$amenities)
df$amenities <- gsub('\\[',"",df$amenities)
df$amenities <- gsub('\\]',"",df$amenities)
df$amenities <- tolower(df$amenities)

# Check most frequent items
out_db <- df %>% 
  mutate(amenities = tolower(amenities))

list_amenities <- map(out_db$amenities, function(x) x %>% strsplit(", ") %>% unlist() ) %>% 
  unlist()

df_amenities <- data.frame(amenities = list_amenities) 

df_viz <- df_amenities %>% 
  count(amenities, name = "frequency") %>% 
  mutate(n_data = nrow(out_db),
         ratio = frequency/nrow(out_db)) %>% 
  arrange(desc(ratio)) %>% 
  head(20)

p <- df_viz %>% 
  ggplot(aes(x = ratio,
             y = amenities %>% reorder(ratio))) +
  geom_col(fill = "#2ca25f") +
  labs(x = "Percentage of Listing",
       y = "Amenities",
       title = "Most Common Amenities") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 12L,
                              face = "bold", hjust = 0.5),
    axis.title.y = element_text(face = "bold"),
    axis.title.x = element_text(face = "bold"),
    legend.position = "none"
  )

p

# Save plot
ggsave(paste0(output, "/top_20_amenities.jpg"), p, width = 6, height = 6)


# Add dummies for the most frequent and sensible amenities 
df[, d_wifi := ifelse(df$amenities %like% "wifi" == T, 1, 0)]
df[, d_long_term_stays_allowed := ifelse(df$amenities %like% "long term stays" == T, 1, 0)]
# df[, d_essentials := ifelse(df$amenities %like% "essentials" == T, 1, 0)]
df[, d_kitchen := ifelse(df$amenities %like% "kitchen" == T, 1, 0)]
# df[, d_hangers := ifelse(df$amenities %like% "hangers" == T, 1, 0)]
# df[, d_hair_dryer := ifelse(df$amenities %like% "hair dryer" == T, 1, 0)]
# df[, d_heating := ifelse(df$amenities %like% "heating" == T, 1, 0)]
df[, d_washer := ifelse(df$amenities %like% "washer" == T, 1, 0)]
# df[, d_iron := ifelse(df$amenities %like% "iron" == T, 1, 0)]
# df[, d_hot_water := ifelse(df$amenities %like% "hot water" == T, 1, 0)]
df[, d_air_conditioning := ifelse(df$amenities %like% "air conditioning" == T, 1, 0)]
df[, d_tv := ifelse(df$amenities %like% "tv" == T, 1, 0)]
# df[, d_shampoo := ifelse(df$amenities %like% "shampoo" == T, 1, 0)]
df[, d_elevator := ifelse(df$amenities %like% "elevator" == T, 1, 0)]
# df[, d_dishes_and_silverware := ifelse(df$amenities %like% "dishes and silverware" == T, 1, 0)]
df[, d_cooking_basics := ifelse(df$amenities %like% "cooking basics" == T, 1, 0)]
# df[, d_refrigerator := ifelse(df$amenities %like% "refrigerator" == T | df$amenities %like% "fridge" == T, 1, 0)]
df[, d_dedicated_workspace := ifelse(df$amenities %like% "dedicated workspace" == T, 1, 0)]
# df[, d_microwave := ifelse(df$amenities %like% "microwave" == T, 1, 0)]
# df[, d_coffee_maker:= ifelse(df$amenities %like% "coffee maker" == T, 1, 0)]

df[, d_breakfast := ifelse(df$amenities %like% "breakfast" == T, 1, 0)]
df[, d_children_friendly := ifelse(df$amenities %like% "children" == T | df$amenities %like% "baby" == T, 1, 0)]
# df[, d_closet := ifelse(df$amenities %like% "closet" == T, 1, 0)]
df[, d_garden := ifelse(df$amenities %like% "garden" == T | df$amenities %like%  "backyard" == T, 1, 0)]
# df[, d_first_aid_kit := ifelse(df$amenities %like% "first aid kit" == T, 1, 0)]
df[, d_balcony := ifelse(df$amenities %like% "balcony" == T, 1, 0)]
df[, d_gym := ifelse(df$amenities %like% "gym" == T, 1, 0)]
df[, d_free_parking := ifelse(df$amenities %like% "free parking" == T, 1, 0)]
df[, d_smoking_allowed := ifelse(df$amenities %like% "smoking allowed" == T, 1, 0)]
# df[, d_lock_on_bedroom_door := ifelse(df$amenities %like% "lock on bedroom" == T, 1, 0)]
df[, d_private_entrance := ifelse(df$amenities %like% "private entrance" == T, 1, 0)]
df[, d_pool := ifelse(df$amenities %like% "pool" == T, 1, 0)]
df[, d_lockbox := ifelse(df$amenities %like% "lockbox" == T, 1, 0)]
# df[, d_indoor_fireplace := ifelse(df$amenities %like% "indoor fireplace" == T, 1, 0)]
df[, d_pets_allowed := ifelse(df$amenities %like% "pets allowed" == T, 1, 0)]
df[, d_luggage_dropoff_allowed := ifelse(df$amenities %like% "luggage dropoff allowed" == T, 1, 0)]
# df[, d_smart_lock := ifelse(df$amenities %like% "smart lock" == T, 1, 0)]

# Frequencies
amenities_vars <- grep("^d_.*", names(df), value=TRUE)
datasummary_skim(df[, amenities_vars, with = FALSE], fmt =  '%.3f' )
# Common features - Above 90%: d_wifi, d_long_term_stays_allowed, d_essentials, d_kitchen, d_washer, d_tv, d_closet
# Rare features - Below 10%: d_indoor_fireplace, d_pets_allowed, d_smart_lock, d_smoking_allowed, d_breakfast, d_garden

# Add total number of amenities
df$n_count_amenities <- str_count(df$amenities, ",")+1
hist(df$n_count_amenities)

```
**Note**: I did not keep all the amenities implied by the barchart as the majority are very common items with more than 90% of the listings having these amenities. Amenities such as hair dryer or hot water were excluded as these items are unlikely to impact apartment prices. Moreover, other items such as having a garden or balcony were included as these are not that common but likely to impact the price of the listings.

### 4. step: Distance from city center
```{r}

# City center: Plaça de Catalunya (lat: 41.38705287353137, lon: 2.1700683272487105)
lat_center <- 41.38705287353137
lon_center <- 2.1700683272487105

# Add distance in km based on the Great-circle distance formula
df$distance_from_center_km <- acos(sin(lat_center*pi/180)*sin(df$latitude*pi/180) + cos(lat_center*pi/180)*cos(df$latitude*pi/180) * cos((df$longitude*pi/180)-(lon_center*pi/180))) * 6371

ggplot(df, aes(distance_from_center_km)) +
  geom_histogram() +
  theme_bw()

# Create distance categories
df$n_distance_from_center_km <- ceiling(df$distance_from_center_km)
df[, .(count = .N), by = .(n_distance_from_center_km)][order(n_distance_from_center_km)] 
datasummary_skim(df$n_distance_from_center_km)

```
**Note**: It was assumed that the center of the city is the Plaça de Catalunya and distance was measured from its geocoordinates.

### 5. Convert property_type to factor
```{r}

# Property type
df$f_property_type <- as.factor(df$property_type)

# Summary
datasummary_skim(df$property_type, 'categorical')

```

### 6. Convert neighborhood_cleansed to factor
```{r}

# Neighbourhood
df$f_neighbourhood_group_cleansed <- as.factor(df$neighbourhood_group_cleansed)

# Summary
datasummary_skim(df$f_neighbourhood_group_cleansed, 'categorical')

```

### 7. step: Add naming convention to the rest of the numeric fields
```{r}

# Numeric
df$n_minimum_nights <- df$minimum_nights
# df$n_maximum_nights <- df$maximum_nights
df$n_bedrooms <- df$bedrooms
df$n_beds <- df$beds
df$n_accommodates <- df$accommodates
df$n_reviews <- df$number_of_reviews
df$n_review_scores_rating <- df$review_scores_rating

```

# Keep only cleaned variables
```{r}

df_final <- df[, .(id)]

# Numeric vars - 11 vars
numeric_vars <- grep("^n_.*", names(df), value=TRUE)
df_numeric <- df[, .SD, .SDcols =  names(df) %in% numeric_vars]
df_final <- cbind(df_final, df_numeric)

# Factors - 3 vars
factor_vars <- grep("^f_.*", names(df), value=TRUE)
df_factor <- df[, .SD, .SDcols =  names(df) %in% factor_vars]
df_final <- cbind(df_final, df_factor)

# Dummies - 27 vars
binary_vars <- grep("^d_.*", names(df), value=TRUE)
df_binary <- df[, .SD, .SDcols =  names(df) %in% binary_vars]
df_final <- cbind(df_final, df_binary)

# Summary - 65 vars
datasummary_skim(df_final, 'numeric')
datasummary_skim(df_final, 'categorical')

```


# Managing missings

```{r}

# Missings
missings <- sapply(df_final, function(x) sum(is.na(x)))
missings[missings > 0]
# n_bathrooms, n_bedrooms, n_beds, n_review_scores_rating,  f_bathrooms 

```

### 1. step: Missing bathrooms
```{r}

# Missing bathrooms - 1 obs. - Add 1 assuming that apartments have at least one bathroom
df_final$flg_n_bathrooms <- ifelse(is.na(df_final$n_bathrooms), 1, 0) # 1 missing
df_final$n_bathrooms <- ifelse(is.na(df_final$n_bathrooms), median(df_final$n_bathrooms, na.rm = T), df_final$n_bathrooms) 
df_final$f_bathrooms <- ifelse(is.na(df_final$f_bathrooms), 2, df_final$f_bathrooms)

# Summary
df_final[, .(count = .N), by = .(n_bathrooms, f_bathrooms, flg_n_bathrooms)][order(flg_n_bathrooms, n_bathrooms)] 

```

### 2. step: Missing bedrooms
```{r}

# Missing bedrooms - 340 obs. - Add 0 assuming that apartments do not have separate bedrooms (studio)
df_final$flg_n_bedrooms <- ifelse(is.na(df_final$n_bedrooms),1, 0) # 340 missing
df_final$n_bedrooms <- ifelse(is.na(df_final$n_bedrooms),0, df_final$n_bedrooms)

# Summary
df_final[, .(count = .N), by = .(n_bedrooms, flg_n_bedrooms)][order(flg_n_bedrooms, n_bedrooms)] 

```

### 3. step: Missing beds
```{r}

# Missing beds - 134 obs. - Add number of accommodates to missing beds
df_final[is.na(n_beds), .(count = .N), by = .(n_accommodates)]

df_final$flg_n_beds <- ifelse(is.na(df_final$n_beds),1, 0)  # 134 missing
df_final$n_beds <-  ifelse(is.na(df_final$n_beds), 
                           df_final$n_accommodates, df_final$n_beds)

# df_final$n_ln_beds <- log(df_final$n_beds)
# df_final$flg_n_ln_beds <- ifelse(is.na(df_final$n_ln_beds),1, 0) # 134 missing
# df_final$n_ln_beds <- ifelse(is.na(df_final$n_ln_beds),0, df_final$n_ln_beds)

# Summary
df_final[, .(count = .N), by = .(n_beds, flg_n_beds)][order(flg_n_beds, n_beds)] 

```

### 4. step: Missing n_review_scores_rating
```{r}

#  Missing n_review_scores_rating - 1631 obs. - Add median 
df_final$flg_n_review_scores_rating <- ifelse(is.na(df_final$n_review_scores_rating),1, 0) # 1631 missing
df_final$n_review_scores_rating <-  ifelse(is.na(df_final$n_review_scores_rating), 
                                           median(df_final$n_review_scores_rating, na.rm = T), df_final$n_review_scores_rating)
# Summary
hist(df_final$n_review_scores_rating)

```
**Note**: Missing rate is high in this variable but it is expected to be an important predictor hence kept for the analysis.

```{r}

# Check missings again
missings <- sapply(df_final, function(x) sum(is.na(x)))
missings[missings > 0] # 0 missings now

```


# Variable categorization

### 1. step: n_reviews
```{r}

# n_reviews: [0], [1-51], [51-100] [100+]
hist(df_final$n_reviews)

df_final$f_reviews <- cut(df_final$n_reviews, c(0,1,51,101,9999), labels=c(0,1,2,3), right = F) 
df_final[, .(count = .N, min = min(n_reviews), max = max(n_reviews), avg_price = mean(n_price)), by = .(f_reviews)]

# Check relationship with price (excl. extremes)
boxplot(n_price ~ f_reviews, data=df_final[n_price < 600])

```

**Note**: Few listings above 100 reviews, group them together.

### 2. step: n_minimum_night
```{r}

# n_minimum_nights: [1],[2-7],[8-14],[14+]
hist(df_final$n_minimum_nights, breaks = 100)

df_final$f_minimum_nights <- cut(df_final$n_minimum_nights, c(1,2,8,9999), labels=c(1,2,3), right = F)
df_final[, .(count = .N, min = min(n_minimum_nights), max = max(n_minimum_nights), avg_price = mean(n_price)), by = .(f_minimum_nights)]

# Check relationship with price (excl. extremes)
boxplot(n_price ~ f_minimum_nights, data=df_final[n_price < 400])

```

**Note**: This variable is most probably not for measuring the minimum nights that the guests should spend at the apartment but the number of available nights. It is better to exclude this variable from later analysis.

### 3. step: Property type
```{r}

# Create categories based on mean price
df_final[, .(count = .N, mean_price = mean(n_price)), by = .(f_property_type)][order(mean_price)] 

df_final <- df_final %>% mutate(
  f_property_type = as.character(f_property_type), 
  f_property_type2 = case_when(
    f_property_type %in% c("Entire rental unit", "Entire home/apt", "Entire residential home", "Entire townhouse", "Casa particular", "Entire place") ~ "Entire home / Apartment",
    f_property_type %in% c("Entire chalet", "Tiny house", "Entire guesthouse", "Entire guest suite") ~ "Entire guesthouse / chalet",
    TRUE ~ f_property_type))

df_final[, .(count = .N, mean_price = mean(n_price)), by = .(f_property_type2)][order(mean_price)] 

# Overwrite the original field
df_final$f_property_type <- df_final$f_property_type2
df_final[,f_property_type2:=NULL]

```

**Note**: Small property type were merged to the bigger groups based on their average prices.


# Final check
```{r}

# Numeric
numeric_vars <- grep("^n_.*", names(df_final), value=TRUE)
df_numeric <- df_final[, .SD, .SDcols =  names(df_final) %in% numeric_vars]
datasummary_skim(df_numeric, fmt =  '%.3f' )

# Dummies
dummy_vars <- grep("^d_.*", names(df_final), value=TRUE)
df_dummy <- df_final[, .SD, .SDcols =  names(df_final) %in% dummy_vars]
datasummary_skim(df_dummy, fmt =  '%.3f' )

# Factors
factor_vars <- grep("^f_.*", names(df_final), value=TRUE)
df_factor <- df_final[, .SD, .SDcols =  names(df_final) %in% factor_vars]
datasummary_skim(df_factor, fmt =  '%.3f' , 'categorical')

```


# Create work file

```{r}

# Drop fields
drops <- c("n_bathrooms", "n_reviews", "n_minimum_nights")
df_final <- df_final[, (drops) := NULL]

# Create work file
write.csv(df_final, file = paste0(data_work,'/airbnb_barcelona_20211207_cleaned.csv'))

```











